Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-03-05 manuscript/climate_sensitivity_draft.Rmd
condensed/ cleaned-up text on identifying drought
#80

We identified droughts within the time period 1950-2009, defining drought [@slette_how_2019] as events with both anomalously dry peak growing season climatic conditions and widespread reductions in tree growth, *i.e.*, droughts that substantially impacted the forest community. We identified three drought years: 1966, 1977, and 1999 (Figs. 1, S2, Table S3). These were the years with the lowest Palmer Drought Severity Index (PDSI) during May-August (MJJA; Table S3), which were identified by @helcoski_growing_2019 as the months of the current year to which annual tree growth was most sensitive at this site. PDSI divisional data for Northern Virginia were obtained from NOAA (https://www7.ncdc.noaa.gov/CDO/CDODivisionalSelect.jsp) in December 2017.

We defined years with widespread growth reduction ("pointer years") as those where >25% of the cored trees experienced >30% reduction in basal area increment ($BAI$) relative to the previous 5 years, following the drought resistance ($Rt$) metric of [@lloret_components_2011]. $Rt$ values <1 indicate growth reductions, whereas values >1.0 indicate above-average growth. Pointer years were identified using the pointRes package [@R-pointRes] in R. Four years met our criteria: 1966, 1977, 1991, and 1999. We excluded 1991 (26.5% of trees experienced >30% growth reduction, mean resistance= -13.8%) because this year was not identified as among the driest of the time period (Table S3). Rather, the severity of growth reduction could probably be explained in part by defoliation by the gypsy moth (*Lymantria dispar* L.), which was documented to have strongly impacted *Quercus* spp. in the area from approximately 1988 through 1995 [@twery_effects_1991].

>>>>

We identified droughts within the time period 1950-2009, defining drought [@slette_how_2019] as events with both anomalously dry peak growing season climatic conditions and widespread reductions in tree growth, *i.e.*, droughts that substantially impacted the forest community. We identified three drought years: 1966, 1977, and 1999 (Figs. 1, S2, Table S3). These were the years with the lowest Palmer Drought Severity Index (PDSI) during May-August (MJJA; Table S3), which were identified by @helcoski_growing_2019 as the months of the current year to which annual tree growth was most sensitive at this site. PDSI divisional data for Northern Virginia were obtained from NOAA (https://www7.ncdc.noaa.gov/CDO/CDODivisionalSelect.jsp) in December 2017. These were also years with widespread tree growth reduction ("pointer years"), here defined as those where >25% of the cored trees experienced >30% reduction in basal area increment ($BAI$) relative to the previous 5 years, following the drought resistance ($Rt$) metric of [@lloret_components_2011]. $Rt$ values <1 and >1 indicate growth reductions and increases, respectively. Pointer years were identified using the pointRes package [@R-pointRes] in R. In addition to the focal drought years, 1991 also met this critera (26.5% of trees experienced >30% growth reduction, mean resistance= -13.8%) but was excluded because it was not among the driest of the time period (Table S3). Rather, the severity of growth reduction could probably be explained in large part by defoliation by gypsy moths (*Lymantria dispar* L.), which was documented to have strongly impacted *Quercus* spp. in the area from approximately 1988 through 1995 [@twery_effects_1991].
